---
title: "Sim fit results"
author: "John Flournoy"
date: "3/13/2020"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstan)
library(tidybayes)
library(tidyr)
library(dplyr)
library(ggplot2)
library(purrr)

knitr::read_chunk('sim_data.R')
```

```{r}
fit_files <- sprintf('fit%02d.csv', 4:7)
gend <- rstan::read_stan_csv('sim.csv')
```

# Generate the data

```{r generate_data, eval = FALSE}
```

# Fits to simulated data

Eight simulated data sets were used to fit the mlmed example (one chain each).  Two data sets timed out (max wall time was 2 days). A few had divergent transitions, and treedepth warnings.

1. Timeout
2. Timeout
3. 3/1000 divergent transitions
4. 1/1000 maximum treedepth (15)
5. No issues
6. No issues
7. No issues
8. 2/1000 divergent transitions; split R-hat > 1.1 warnings for 2 parameters

# Example fit

## ROI random effects

```{r}
params <- tidybayes::gather_draws(gend, `(.*_(a|b|cp|dy|dm|ty|tm).*|(a|b|cp|dy|dm|ty|tm))`, regex = TRUE)
cor_params <- tidybayes::gather_draws(gend, `.*Omega.*`, regex = TRUE)
tau_params <- tidybayes::gather_draws(gend, `.*Tau.*`, regex = TRUE)

fit <- rstan::read_stan_csv(fit_files[[1]])
fitnum <- as.numeric(gsub('fit(\\d{2}).csv', '\\1', fit_files[[1]]))

params_from_sim <- tidybayes::gather_draws(fit, `(.*_(a|b|cp|dy|dm|ty|tm).*|(a|b|cp|dy|dm|ty|tm))`, regex = TRUE)

cor_params_from_sim <- tidybayes::gather_draws(fit, `.*Omega.*`, regex = TRUE)
tau_params_from_sim <- tidybayes::gather_draws(fit, `.*Tau.*`, regex = TRUE)

v_pars <- params_from_sim %>% group_by(.variable) %>%
  summarize(mean(.value)) %>%
  left_join(params[params$.iteration == fitnum,]) %>%
  filter(grepl('v_', .variable)) %>%
  extract(.variable, c('var', 'var2', 'idx'), regex = '(v_)(.*)\\.(.*)')
```

```{r}
v_pars %>%
  ggplot(aes(x = .value, y = `mean(.value)`)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(~var2, scales = 'free') +
  geom_abline(intercept = 0, slope = 1)
```

## ID random effects

```{r}
  params_from_sim %>% group_by(.variable) %>%
    summarize(mean(.value)) %>%
    left_join(params[params$.iteration == fitnum,]) %>%
    filter(grepl('^u_', .variable)) %>%
    extract(.variable, c('var', 'var2', 'idx'), regex = '(u_)(.*)\\.(.*)') %>%
    ggplot(aes(x = .value, y = `mean(.value)`)) +
    geom_point() +
    geom_smooth(method = 'lm') +
    facet_wrap(~var2, scales = 'free') +
    geom_abline(intercept = 0, slope = 1)
```

## Population parameters

```{r}
params_from_sim %>% group_by(.variable) %>%
  summarize(mean(.value)) %>%
  left_join(params[params$.iteration == fitnum,]) %>%
  filter(!grepl('^(v|u|tau)_', .variable)) %>%
  ggplot(aes(x = .value, y = `mean(.value)`)) +
  geom_abline(intercept = 0, slope = 1) +
  geom_point() +
  geom_smooth(method = 'lm')
```

## RE Variances

```{r}
params_from_sim %>% group_by(.variable) %>%
  summarize(mean(.value)) %>%
  left_join(params[params$.iteration == fitnum,]) %>%
  filter(grepl('tau', .variable)) %>%
  ggplot(aes(x = .value, y = `mean(.value)`)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

# All fits to simulated data

Plots have generating value on the X axis, and sample medians, along with 95% credible intervals, on the Y axis.

```{r}
#https://tbradley1013.github.io/2018/10/01/calculating-quantiles-for-groups-with-dplyr-summarize-and-purrr-partial/
p <- c(.025, .5, .975)
p_names <- map_chr(p, ~paste0(.x*100, "%"))

p_funs <- map(p, ~partial(quantile, probs = .x, na.rm = TRUE)) %>% 
  set_names(nm = p_names)

simfitpars_ <- lapply(fit_files, function(fit_file){
  fit <- rstan::read_stan_csv(fit_file)
  fitnum <- as.numeric(gsub('fit(\\d{2}).csv', '\\1', fit_file))
  
  params_from_sim <- tidybayes::gather_draws(fit,
                                             `(.*_(a|b|cp|dy|dm|ty|tm).*|(a|b|cp|dy|dm|ty|tm)|.*Tau.*)`, 
                                             regex = TRUE) %>%
    group_by(.variable) %>%
    summarize_at(.vars = vars(.value), .funs = p_funs) %>%
    mutate(.iteration = fitnum)
  return(params_from_sim)
})

simfitpars <- bind_rows(simfitpars_)

gen_and_simfit_pars <- left_join(simfitpars, params, by = c('.variable', '.iteration'))
```

## ROI-varrying parametrs (V)

```{r fig.width=12, fig.height=12}
gen_and_simfit_pars %>%
  filter(grepl('v_', .variable)) %>%
  extract(.variable, c('var', 'var2', 'idx'), regex = '(v_)(.*)\\.(.*)') %>%
  ggplot(aes(x = .value, y = `50%`)) +
  geom_segment(aes(y = `2.5%`, yend = `97.5%`, xend = .value), alpha = .5) + 
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(.iteration~var2, scales = 'free', ncol = 7) +
  geom_abline(intercept = 0, slope = 1) + theme_minimal()
```

## ID-varrying parametrs (U)

```{r fig.width=12, fig.height=12}
gen_and_simfit_pars %>%
  filter(grepl('^u_', .variable)) %>%
  extract(.variable, c('var', 'var2', 'idx'), regex = '(u_)(.*)\\.(.*)') %>%
  ggplot(aes(x = .value, y = `50%`)) +
  geom_segment(aes(y = `2.5%`, yend = `97.5%`, xend = .value), alpha = .5) + 
  geom_point() +
  geom_smooth(method = 'lm') +
  facet_wrap(.iteration~var2, scales = 'free', ncol = 7) +
  geom_abline(intercept = 0, slope = 1) + theme_minimal()
```

## RE Variance parameters

```{r fig.width=10, fig.height=5}
gen_and_simfit_pars %>%
  filter(grepl('tau', .variable)) %>%
  extract(.variable, c('var', 'group', 'par'), regex = '(tau)_(.*)_(.*)') %>%
  ggplot(aes(x = .value, y = `50%`)) +
  geom_abline(intercept = 0, slope = 1, alpha = .5) + 
  geom_segment(aes(y = `2.5%`, yend = `97.5%`, xend = .value), alpha = .5) +
  geom_point() +
  facet_wrap(group~par, scales = 'free', ncol = 7) +
  theme_minimal()
```